---
title: "Practical Machine Learning Course project"
author: "Kirill Shmidt"
date: '29 февраля 2016 г '
output: html_document
---

```{r setup, include=FALSE}
library(caret)
load("envi.R.RData")
knitr::opts_chunk$set(echo = TRUE)
```

#Abstract

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. In this project, I use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of your project is to predict the manner in which they did the exercise.

#Results
I got 20/20 correct results with gbm prediction model.
Results of modeling:
```{r echo=FALSE}

finalpred

```

#Data summary

I have data sets diveded in two subsets:

*Training - 19216 observations
(Testing - 20 observation. below i call it Final Testing.

General goal of this work is prediction of exrecise performed in testing set

Sets consist of 160 variables:

1. Index

2. name of person, who performed observation

3-5. Time stamps and window information

6-159. Data from accelerometers

160. Performed action in observation

#Data preprocessing

1. Clearing - many colomn have NA or blank data. I dicided to delete this colomns. 

2. First 7 colomns doesnt carry any neccesary information about repforming action

3. Rows with variable "newwindow" = "yes" is summary of other rows. Our testing set doesn't have rows with variable "newwindow" = "yes", so i delete this rows.

Colomns in analysys after clearing: 
8:11, 37:49, 60:68, 84:86, 102, 113:124, 140,151:160

4. For estimation of out of sample error training set was subseted in two groups: training subset(80% of observations), testing subset (20% of observations)

```{r echo=F}
library(caret)
load("envi.R.RData")
```

```{r echo=TRUE, eval=F}
tra <- training[training$new_window=="no",c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140,151:160)]
tra[,1:52] <- sapply(tra[,1:52], FUN = as.character)
tra[,1:52] <- sapply(tra[,1:52], FUN = as.numeric)
tes <- testing[,c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140,151:160)]
tes[,1:52] <- sapply(tes[,1:52], FUN = as.character)
tes[,1:52] <- sapply(tes[,1:52], FUN = as.numeric)

inTrain <- createDataPartition(tra[,53], p=0.8, list=F)
tra_training <- tra[inTrain,]
tra_testing <- tra[-inTrain,]
```


#Prediction models
Used methods with caret package with defaults options:

1. rpart

2. lda

3. gbm

#Results of modeling

I perform prediction of testing subset using models builded on training subset. I diceded to use accuracy as measure of out of sample error.
Best model should be used on Final Testing data set.

*rpart
```{r echo=FALSE}

confusionMatrix(tra_testing$classe, predTree)
```

*lda
```{r echo=FALSE}

confusionMatrix(tra_testing$classe, predlda)
```

*gbm
```{r echo=FALSE}
confusionMatrix(tra_testing$classe, predgbm)
```

gbm have the best accuracy. This model was used in Final Testing.

#Final testing

I got 20/20 correct results with gbm prediction model.
Results of model:
```{r echo=FALSE}

finalpred

```

#Appendix

Full code

```{r echo=TRUE, eval=F}

library(caret)

# readind data
training <-read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

# delete unnesessaity variable - id, names, time stamps, etc
# maybe delete all data with NA and with blank fields?
#Clear off window = YES becose we dont have such data in testing


tra <- training[training$new_window=="no",c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140,151:160)]
tra[,1:52] <- sapply(tra[,1:52], FUN = as.character)
tra[,1:52] <- sapply(tra[,1:52], FUN = as.numeric)
tes <- testing[,c(8:11, 37:49, 60:68, 84:86, 102, 113:124, 140,151:160)]
tes[,1:52] <- sapply(tes[,1:52], FUN = as.character)
tes[,1:52] <- sapply(tes[,1:52], FUN = as.numeric)

#cross validation. Main idea - create 2 subsets of "training data" witch can be used for OOS measure
#Final testing set would be our final check, after ill find best model
inTrain <- createDataPartition(tra[,53], p=0.8, list=F)
tra_training <- tra[inTrain,]
tra_testing <- tra[-inTrain,]

#For parralel calcalation
cores <- 4
cl <- makeCluster(cores)
registerDoParallel(cores)

modelTree <- train(tra_training$classe~., data=tra_training, method="rpart")

modellda <- train(tra_training$classe~., data=tra_training, method="lda")

modelgbm <- train(tra_training$classe~., data=tra_training, method="gbm")

stopCluster(cl)

predTree <- predict(modelTree,tra_testing)
predlda <- predict(modellda, tra_testing)
predgbm <- predict(modelgbm, tra_testing)

confusionMatrix(tra_testing$classe, predTree)
confusionMatrix(tra_testing$classe, predlda)
confusionMatrix(tra_testing$classe, predgbm)

finalpred <- predict(modelgbm, tes)
```
